---
title: "Barbell Lifts Analysis"
output: html_document
---

#Synopsis
The goal of this project is to build a model with data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict the manner in which they did an exercise - barbell lifts.
The model should should predict "classe" variable with any other variables.

#Preparing data and dependencies
Initial preparations - loading required packages, loading data.

```{r}
suppressMessages(library(caret))

# Load data
srcTrain <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
if(!file.exists("pml-training.csv")){
  download.file(srcTrain, destfile = "pml-training.csv", method = "curl", mode="wb")
}

srcTest <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
if(!file.exists("pml-testing.csv")){
  download.file(srcTest, destfile = "pml-testing.csv", method = "curl", mode="wb")
}
pml.training <- read.csv("pml-training.csv")
pml.testing <- read.csv("pml-testing.csv")
```

##Subseting and cleaning data
Some variables are for statistical purpose only. Because they are derived from the others (also correlated), than can be removed to simplify building prediction model.

```{r}
remove_cols_reqex <- "^(kurtosis|skewness|min|max|stddev|total|var|avg|ampl|num_window|cvtd_timestamp|X|new_window)"
pml.training <- pml.training[,grep(remove_cols_reqex, names(pml.training),invert=T)]
```

Split training data into training and validation subsets
```{r}
sub <- createDataPartition(y=pml.training$classe, p=.7, list=F) 
training <- pml.training[sub,]
validation <- pml.training[-sub,]
```

##Machine learning

Cross-validation with 3 fold, and allow running for paralell for shorten time
```{r}
ctrl <- trainControl(method="cv",number=3,allowParallel=T)
```

User random forest to build a model.
```{r cache=TRUE}
modelRF <- train(classe~., data=training, method="rf", trControl=ctrl, allowParalell=T)
modelRF
```

Build confusion matrix for validation data
```{r}
confusionMatrix(predict(modelRF,validation),validation$classe)
```

The random forest accuracy on training set is 99.75% on the validation set. It suggests, that the out-of-sample error could be small.

##Test
Predict results on test data

```{r}
pml.testing.predicted = predict(modelRF,pml.testing)
pml.testing.predicted
```

```{r echo=FALSE}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(pml.testing.predicted)
```



